---
title: "Practice Machine Learning Assignment"
author: "Mantas Traigys"
date: "Friday, October 24, 2014"
output: html_document
---
### Introduction
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

### Loading and preprocessing the data

```{r}
download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "C:/Users/Mantas/Documents/assignment9/pml-training.csv")
download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "C:/Users/Mantas/Documents/assignment9/pml-testing.csv")
train <- read.csv("./pml-training.csv", na.string = c("NA", ""))
test <-  read.csv("./pml-testing.csv", na.string = c("NA", ""))
```

### Splitting the data

```{r}
library(caret)
set.seed(852)
trainset <- createDataPartition(train$classe, p = 0.75, list = FALSE)
Training <- train[trainset, ]
CrV <- train[-trainset, ]
```

### Cleaning the data 
To clean dataset for futher analysis and prediction I used near Zero variance. Also I cleared variables, which can corrupt prediction model (such as user_name). Moreover, I subsetted columns, which contains missing values.
```{r}
zero <- nearZeroVar(Training)
Training <- Training[, -zero]
unusable <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", 
                 "cvtd_timestamp", "num_window")
Training <- Training[, !names(Training) %in% unusable]
Usable <- c((colSums(!is.na(Training[,-ncol(Training)])) >= 1*nrow(Training)))
Training   <-  Training[, Usable]

```
 
### Prediction model
I used Random forest method to build my model, because it is accurated and my dataset is not big enough to make any problems of running RandomForest. 
```{r}
library(randomForest)
Modelfit <- randomForest(classe ~ ., data = Training, importance = TRUE, ntrees = 12)
preCrV <- predict(Modelfit, CrV)
print(confusionMatrix(preCrV, CrV$classe))
testpred <- predict(Modelfit, test)
testpred
```
